Model Comparison

### 2020-07-14 Compare Models.Rmd
Updates to the paper, and generating a standardized patient/sample database for metadata

kb.perf = do.call(bind_rows, lapply(list.files(data,full.names = T), function(f) {
  load(paste0(f,'/all.pt.alpha.Rdata'),verbose=F)

  nonz = names(which(sapply(performance.at.1se, nrow) > 0))
# do.call(bind_rows,lapply(coefs, function(df){
#   tibble(n=nrow(df), mean.stability=mean(rowSums(df[,-1])/50), sd.stability=sd(rowSums(df[,-1])/50) )
# })) %>% mutate(alpha=names(coefs), kb=basename(f))

  do.call(bind_rows, performance.at.1se) %>% 
    mutate(alpha=nonz, kb = basename(f), n.coefs=sapply(coefs[nonz], nrow)) %>% 
    dplyr::select(mean, sme, `lambda-1se`, alpha, n.coefs, kb)

@@

@@

```{r fig.height=7, fig.width=8}
rc9 = do.call(bind_rows, lapply(list.files(data,full.names = T), function(f) {
  load(paste0(f,'/loo_0.9.Rdata'),verbose=F)
  roc = pROC::roc(Status ~  Prediction, data=pg.samp, auc=T, ci=T, of='thresholds',transpose=T, quiet=T)
  pROC::coords(roc, "best", transpose=F) %>% mutate( auc = as.numeric(pROC::auc(roc)), kb = basename(f) )
}))

rc9 = rc9 %>% set_names(c('Threshold','FPR','TPR','AUC','kb')) %>% 
  mutate(kb = factor(kb,levels=c('15kb','50kb','100kb','500kb'), ordered=T), FPR = 1-FPR) %>% arrange(kb) %>%
  dplyr::select(-Threshold)

p = ggplot(melt(rc9, id.vars='kb'), aes(variable, value, group=kb)) + ylim(0,1) + 
  geom_bar(aes(fill=kb), stat='identity', position='dodge') + 
  geom_text(aes(label=round(value,2)), position=position_dodge(width=1)) + 
  scale_fill_brewer(palette='Set1') + 
  labs(title='Discovery ROC', subtitle = 'alpha=0.9', x='', y='%') + 
  theme_minimal() + theme(text=element_text(size=12))
p
ggsave(filename='plots/parameters/training_kb_auc.png', plot = p, width = 6, height = 6, units = 'in', dpi = 300)

@@

val.preds = purrr::map(list.files(vdir, full.names=T), function(dir) {
  do.call(bind_rows, purrr::map(list.files(paste0(dir,'/predictions_5e6_all/0.9'), 'predictions.tsv', full.names = T, recursive = T), function(f) {
    pd = read_tsv(f, col_types = 'cddcccDcccc') %>% dplyr::select(`Hospital Research ID`, everything())
    left_join(pd, val.info %>% dplyr::select(`Hospital Research ID`, Status, Samplename, Cohort), by=c('Hospital Research ID', 'Sample'='Samplename')) %>% 
      mutate(kb = basename(dir)) %>% mutate(Risk = factor(Risk, levels=c('Low','Moderate','High'), ordered=T)) 
  }))
})
names(val.preds) = list.files(vdir)

val.auc = do.call(bind_rows, purrr::map(val.preds, function(kb) {
  roc = pROC::roc(Status ~ Probability, data=kb, auc=T, ci=T, of='thresholds',transpose=T, quiet=T)
  pROC::coords(roc, "best", transpose=F) %>% mutate( auc = as.numeric(pROC::auc(roc)), kb = unique(kb$kb) )
}))

val.auc = val.auc %>% set_names(c('Threshold','FPR','TPR','AUC','kb')) %>% 
  mutate(kb = factor(kb,levels=c('15kb','50kb','100kb','500kb'), ordered=T), FPR = 1-FPR) %>% arrange(kb) %>%
  dplyr::select(-Threshold)

ggplot(melt(val.auc, id.vars='kb'), aes(variable, value, group=kb)) + ylim(0,1) +
  geom_bar(aes(fill=kb), stat='identity', position='dodge') + 
  geom_text(aes(label=round(value,2)), position=position_dodge(width=1)) + 
  scale_fill_brewer(palette='Set1') + 
  labs(title='Validation ROC (alpha=0.9)', x='', y='%') + theme_minimal()

@@

```{r fig.height=7, fig.width=8}
aucs = bind_rows(rc9 %>% mutate(Cohort = 'Discovery'), val.auc %>% mutate(Cohort = 'Validation'))

p = ggplot(aucs, aes(kb, AUC, group=Cohort)) + ylim(0,1) + 
  geom_bar(aes(fill=Cohort), stat='identity', position='dodge') + 
  geom_text(aes(label=round(AUC,2)), position=position_dodge(width=1)) + 
  scale_fill_brewer(palette='Set1') + 
  labs(title='AUC discovery & validation', x='', y='AUC') + 
  theme_minimal() + theme(text = element_text(size=12))
p
ggsave(filename='plots/parameters/tv_auc_kb_compare.png', plot = p, width = 6, height = 6, units = 'in', dpi = 300)

@@

```{r, fig.width=10,fig.height=7, eval=F}
## 0.5
loo.pred.confidence = purrr::map( list.files(data,full.names = T), function(f) {
  #print(f)
  load(paste0(f,'/loo_0.5.Rdata'),verbose=F)
  suppressWarnings(pred.conf(pg.samp)) %>% mutate(kb = basename(f))
}) 
names(loo.pred.confidence) = basename(list.files(data,full.names = T))

@@

```{r, fig.width=8, fig.height=12}
## ALL LOO is at alpha=0.9
loo.pred.confidence = purrr::map( list.files(data,full.names = T), function(f) {
  #print(f)
  load(paste0(f,'/loo_0.9.Rdata'),verbose=F)
  suppressWarnings(pred.conf(pg.samp)) %>% mutate(kb = basename(f))
}) 
names(loo.pred.confidence) = list.files(data)

@@

per.samp.train = lapply(list.files(data,full.names = T), function(f) {
  load(paste0(f,'/loo_0.9.Rdata'),verbose=F)

@@

  pg.samp %>% group_by(Status, Risk) %>% tally %>% spread(Risk,n) %>% 
    left_join(pg.samp %>% group_by(Status) %>% tally, by='Status') %>%
    dplyr::rename('Total' = 'n') %>%
    mutate_at(vars(Low,Moderate,High), funs(./Total)) %>% mutate(kb = kb)
})
names(per.samp.train) = list.files(data)

@@

@@

```{r, fig.width=8, fig.height=12, eval=F}
test.preds.per.kb09 = purrr::map(list.files(data,full.names = T), function(f) {
  read_tsv(list.files(f, 'test_patients_preds_0.9.tsv', full.names = T), col_types = 'cdccdccclddccccccddccccdddccdddcdd') %>% dplyr::select(Hospital.Research.ID, Status, Endoscopy.Year, Pathology, Prediction, RR, Samplename, Path.ID) %>% mutate(kb = basename(f))
}) 
names(test.preds.per.kb09) = basename(list.files(data))

test.preds.per.kb09 = lapply(test.preds.per.kb09, function(kb) {
  kb %>% mutate(quants = cut(Prediction, breaks=seq(0,1,0.1), include.lowest = T)) %>% 
    left_join(loo.pred.confidence[[unique(kb$kb)]] %>% dplyr::select(quants, Risk), by='quants') %>%
    mutate(Risk = factor(Risk, levels=c('Low','Moderate','High'), ordered=T)) 
})

test.samples = all.patient.info %>% filter(Set == 'Test') %>% group_by(Status) %>% tally

test.counts09 = lapply(test.preds.per.kb09, function(kb) {
  kb %>% group_by(Status, Risk) %>% tally %>% spread(Risk,n) %>% ungroup %>%
    mutate(kb = unique(kb$kb)) %>% left_join(test.samples, by='Status') %>% 
    mutate_if(is.numeric, funs(./n))
})

plist = lapply(test.counts09, function(kb) {
  tt = unique(kb$kb)
  kb = kb  %>% ungroup %>% dplyr::select(-kb,-n)

  ggplot(melt(kb, id.vars=c('Status')), aes(variable, value, group=Status, fill=Status)) + ylim(0,1) + 
    geom_bar(stat='identity', position='dodge') + geom_text(aes(label=round(value,2)), position = position_dodge(width=1)) +
    #scale_fill_manual(values=c('#4575B4','#A50026'),labels=c('NP','P')) + 
    labs(title=tt, x='', y='sample ratio') + theme_minimal()
})

do.call(grid.arrange, c(plist[c('50kb','100kb','15kb','500kb')], top='Test set predictions, alpha=0.9'))
```

@@

>>>>

kb.dirnames = grep('kb', list.files(data,full.names = T), value=T)

kb.perf = do.call(bind_rows, lapply(kb.dirnames, function(f) {
  load(paste0(f,'/all.pt.alpha.Rdata'),verbose=F)
  nonz = names(which(sapply(performance.at.1se, nrow) > 0))
# do.call(bind_rows,lapply(coefs, function(df){
#   tibble(n=nrow(df), mean.stability=mean(rowSums(df[,-1])/50), sd.stability=sd(rowSums(df[,-1])/50) )
# })) %>% mutate(alpha=names(coefs), kb=basename(f))
  do.call(bind_rows, performance.at.1se) %>% 
    mutate(alpha=nonz, kb = basename(f), n.coefs=sapply(coefs[nonz], nrow)) %>% 
    dplyr::select(mean, sme, `lambda-1se`, alpha, n.coefs, kb)

@@

kb.perf %>% dplyr::select(-`lambda-1se`) %>% write_tsv('plots/exfig9a.tsv')

@@

```{r fig.height=7, fig.width=8}
threshold = 'best'
rc9 = do.call(bind_rows, lapply(kb.dirnames, function(f) {
  load(paste0(f,'/loo_0.9.Rdata'),verbose=F)
  roc = pROC::roc(Status ~  Prediction, data=pg.samp, auc=T, ci=T, of='thresholds',quiet=T)

  bind_cols(
    as_tibble(pROC::ci.auc(roc)) %>% mutate(m = c('auc.low', 'auc', 'auc.high')) %>% spread(m,value),
    as_tibble(data.frame(pROC::ci.coords(roc, threshold, transpose=F))) %>% dplyr::select(-matches('threshold')) %>% 
      set_names(c('specificity.low', 'specificity', 'specificity.high', 'sensitivity.low', 'sensitivity', 'sensitivity.high')) ) %>%
    mutate(kb = basename(f))
}))

rc9 = rc9 %>%  mutate(kb = factor(kb,levels=c('15kb','50kb','100kb','500kb'), ordered=T)) %>% arrange(kb)

df = rc9 %>% dplyr::select(matches('auc|kb'))
pos = position_dodge(width=0.9)
p = ggplot() + 
  geom_bar(data=df, aes(x=kb, y=auc, group=kb, fill=kb), stat='identity', position = pos, show.legend = F) + 
  geom_errorbar(data=df, aes(x=kb, ymin=auc.low, ymax=auc.high, group=kb), position = pos, width=0.2, color='grey39') +
  scale_fill_brewer(palette='Set1') + 
  scale_y_continuous(expand=c(0,0), limits=c(0,1)) + 
  labs(title='', subtitle = 'alpha=0.9', x='AUC', y='%') + 
  theme_minimal() + theme(text=element_text(size=12), axis.text.x = element_text(angle=45, hjust=1))

df = rc9 %>% dplyr::select(matches('sensitivity|kb'))
p1 = ggplot() + 
  geom_bar(data=df, aes(x=kb, y=sensitivity, group=kb, fill=kb), stat='identity', position = pos, show.legend = F) + 
  geom_errorbar(data=df, aes(x=kb, ymin=sensitivity.low, ymax=sensitivity.high, group=kb), position = pos, width=0.2, color='grey39') +
  scale_fill_brewer(palette='Set1') + 
  scale_y_continuous(expand=c(0,0), limits=c(0,1)) + 
  labs(title='', subtitle = '', x='TPR', y='') + 
  theme_minimal() + theme(text=element_text(size=12), axis.text.x = element_text(angle=45, hjust=1), axis.text.y = element_blank())

df = rc9 %>% dplyr::select(matches('specificity|kb'))
p2 = ggplot() + 
  geom_bar(data=df, aes(x=kb, y=1-specificity, group=kb, fill=kb), stat='identity', position = pos, show.legend = F) + 
  geom_errorbar(data=df, aes(x=kb, ymin=1-specificity.low, ymax=1-specificity.high, group=kb), position = pos, width=0.2, color='grey39') +
  scale_fill_brewer(palette='Set1') + 
  scale_y_continuous(expand=c(0,0), limits=c(0,1)) + 
  labs(title='', subtitle = '', x='FPR', y='') + 
  theme_minimal() + theme(text=element_text(size=12), axis.text.x = element_text(angle=45, hjust=1), axis.text.y = element_blank())

rc9 %>% dplyr::rename('auc.ci.high'='auc.high', 'auc.ci.low'='auc.low') %>% write_tsv('plots/exfig9b.tsv')

grid.arrange(p,p1,p2,nrow=1)
ggsave(filename='plots/parameters/training_kb_auc.png', plot = grid.arrange(p,p1,p2,nrow=1,top='Discovery ROC'), width = 6, height = 6, units = 'in', dpi = 300)
# rc9 = rc9 %>% set_names(c('Threshold','FPR','TPR','AUC','AUC.CI.Low','AUC.CI.High','kb')) %>% 
#   mutate(kb = factor(kb,levels=c('15kb','50kb','100kb','500kb'), ordered=T), FPR = 1-FPR) %>% arrange(kb) %>%
#   dplyr::select(-Threshold)
```

@@

val.preds = purrr::map(list.files(vdir, full.names=T), function(dir) {
  print(dir)
  do.call(bind_rows, purrr::map(list.files(paste0(dir,'/predictions_5e6_all/0.9'), 'predictions.tsv', full.names = T, recursive = T), function(f) {
    pd = read_tsv(f, col_types = 'cddcccDcccc') %>% dplyr::select(`Hospital Research ID`, everything())
    left_join(pd, val.info %>% dplyr::select(`Hospital Research ID`, Status, Samplename, Cohort), by=c('Hospital Research ID', 'Sample'='Samplename')) %>% 
      mutate(kb = basename(dir)) %>% mutate(Risk = factor(Risk, levels=c('Low','Moderate','High'), ordered=T)) 
  }))
})
names(val.preds) = list.files(vdir)

val.auc = do.call(bind_rows, purrr::map(val.preds, function(kb) {
  roc = pROC::roc(Status ~ Probability, data=kb, auc=T, ci=T, of='thresholds', quiet=T)
  #pROC::coords(roc, "best", transpose=F) %>% mutate( auc = as.numeric(pROC::auc(roc)), kb = unique(kb$kb) )

  bind_cols(
    as_tibble(pROC::ci.auc(roc)) %>% mutate(m = c('auc.low', 'auc', 'auc.high')) %>% spread(m,value),
    as_tibble(data.frame(pROC::ci.coords(roc, threshold, transpose=F))) %>% dplyr::select(-matches('threshold')) %>% 
      set_names(c('specificity.low', 'specificity', 'specificity.high', 'sensitivity.low', 'sensitivity', 'sensitivity.high')) ) %>%
    mutate(kb = unique(kb$kb))
}))

# val.auc = val.auc %>% set_names(c('Threshold','FPR','TPR','AUC','kb')) %>% 
#   mutate(kb = factor(kb,levels=c('15kb','50kb','100kb','500kb'), ordered=T), FPR = 1-FPR) %>% arrange(kb) %>%
#   dplyr::select(-Threshold)
# 
# ggplot(melt(val.auc, id.vars='kb'), aes(variable, value, group=kb)) + ylim(0,1) +
#   geom_bar(aes(fill=kb), stat='identity', position='dodge') + 
#   geom_text(aes(label=round(value,2)), position=position_dodge(width=1)) + 
#   scale_fill_brewer(palette='Set1') + 
#   labs(title='Validation ROC (alpha=0.9)', x='', y='%') + theme_minimal()
```

@@

```{r fig.height=7, fig.width=8}
aucs = bind_rows(rc9 %>% mutate(Cohort = 'Discovery'), val.auc %>% mutate(Cohort = 'Validation'))
aucs = aucs %>% dplyr::select(matches('auc|kb|Cohort')) %>% dplyr::rename('AUC'='auc') %>% 
  mutate(kb = factor(kb, levels=c('15kb','50kb','100kb','500kb'), ordered=T))

aucs %>% dplyr::rename('AUC.CI.high'= 'auc.high', 'AUC.CI.low'='auc.low')  %>% write_tsv('plots/exfig9c.tsv')

p = ggplot(aucs, aes(kb, AUC, group=Cohort)) + ylim(0,1) + 
  geom_bar(aes(fill=Cohort), stat='identity', position=pos) + 
  geom_errorbar(aes(ymin=auc.low, ymax=auc.high), width=0.2, color='grey39', stat='identity', position=pos) +
  geom_text(aes(label=round(AUC,2), y=auc.low-0.05), position=position_dodge(width=1)) + 
  scale_fill_brewer(palette='Set1') + 
  labs(title='AUC discovery & validation', x='', y='AUC') + 
  theme_minimal() + theme(text = element_text(size=12))
p
ggsave(filename='plots/parameters/tv_auc_kb_compare.png', plot = p, width = 6, height = 6, units = 'in', dpi = 300)

@@

```{r, fig.width=10,fig.height=7, eval=F}
## 0.5
loo.pred.confidence = purrr::map( list.files(data,full.names = T), function(f) {
  print(f)
  load(paste0(f,'/loo_0.5.Rdata'),verbose=F)
  suppressWarnings(pred.conf(pg.samp)) %>% mutate(kb = basename(f))
}) 
names(loo.pred.confidence) = basename(list.files(data,full.names = T))

@@

```{r, fig.width=8, fig.height=12}
files = grep('supp', list.files(data,full.names = T), invert = T, value = T)
## ALL LOO is at alpha=0.9
loo.pred.confidence = purrr::map( files, function(f) {
  #print(f)
  load(paste0(f,'/loo_0.9.Rdata'),verbose=F)
  suppressWarnings(pred.conf(pg.samp)) %>% mutate(kb = basename(f))
}) 
names(loo.pred.confidence) = basename(files)

@@

lapply(files, function(f) {
  load(paste0(f,'/loo_0.9.Rdata'),verbose=F)

@@

  pg.samp %>% group_by(Status, Risk) %>% tally %>% spread(Risk,n) %>% 
    left_join(pg.samp %>% group_by(Status) %>% tally, by='Status') %>%
    dplyr::rename('Total' = 'n') %>% mutate(kb = kb)

}) %>% bind_rows() %>% dplyr::rename('Total samples' = 'Total') %>% write_tsv('plots/exfig10a.tsv')

per.samp.train = lapply(files, function(f) {
  load(paste0(f,'/loo_0.9.Rdata'),verbose=F)

  pg.samp %>% group_by(Status, Risk) %>% tally %>% spread(Risk,n) %>% 
    left_join(pg.samp %>% group_by(Status) %>% tally, by='Status') %>%
    dplyr::rename('Total' = 'n') %>%
    mutate_at(vars(Low,Moderate,High), funs(./Total)) %>% mutate(kb = kb)
})
names(per.samp.train) = basename(files)

@@

grid.arrange(p)

@@

@@

lapply(val.preds, function(kb) {
  totals = kb %>% dplyr::filter(Cohort != 'Old progressor scrolls' & Pathology != 'OAC' ) %>% group_by(Status) %>% tally

  kb %>% dplyr::filter(Cohort != 'Old progressor scrolls' & Pathology != 'OAC')  %>% 
    group_by(Status, Risk) %>% tally %>% spread(Risk,n) %>% ungroup %>%
    mutate(kb = unique(kb$kb)) %>% left_join(totals, by='Status')
}) %>% bind_rows() %>% dplyr::rename('Total samples' = 'n') %>% dplyr::select(kb, Status, everything()) %>% write_tsv('plots/exfig10b.tsv')
